### 2023-01-25 paper/manuscript.Rmd
prep new submission, include Tong re-write

---
title: "Global tradeoffs between prosperity, inequality, and the environment"
author: |
    | \small Tong Wu^1,§^, Juan C. Rocha^2,3,§^, Kevin Berry^4^, Tomas Chaigneau^5^, Emilie Lindkvist^2^, Maike Hamann^6^, Jiangxiao Qiu^7^, 
    | \small Caroline Schill^2,8^, Alon Shepon^9^, Anne-Sophie Crépin^8^, and Carl Folke^2,8^
    | \small
    | \footnotesize ^1^The Natural Capital Project and Stanford Center on China's Economy and Institutions, Freeman Spogli 
    | \footnotesize Institute for International Studies, Stanford University
    | \footnotesize ^2^Stockholm Resilience Centre, Stockholm University
    | \footnotesize ^3^Future Earth
    | \footnotesize ^4^College of Business and Public Policy, University of Alaska Anchorage
    | \footnotesize ^5^Environment and Sustainability Institute, University of Exeter
    | \footnotesize ^6^Center for Complex Systems in Transition, Stellenbosch University
    | \footnotesize ^7^School of Forest Resources & Conservation, Fort Lauderdale Research and Education Center, University of Florida
    | \footnotesize ^8^Beijer Institute of Ecological Economics, The Royal Swedish Academy of Sciences
    | \footnotesize ^9^Department of Environmental Studies, The Porter School of the Environment and Earth Science, Tel Aviv University
    | \footnotesize ^§^These authors made equal contributions
#   | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}

output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
csl: /Users/juanrocha/Documents/styles/apa.csl
bibliography: references_ei.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

```{r data, include=FALSE}
library(tidyverse)
library(here)

## word counts
## Abstract 150
## Intro 697
## Method 957
## Results 621
## Discussion 1434
## Conclusion 267
## Total main 3976 + 150 abstract
## 
## Targets: 
# 
# - NatComms: 5000w | 150w abstract | 4 figs or tables | Methods end 3000w | Refs
# 70max
# 
# - WorldDev: | 4500-10k words including all (refs, captions, tables, etc) | 300w
# abstract
```

\doublespacing

**Author contributions:**  Conceptualization was a team effort, TW and JCR produced the first manuscript, TW, JCR and JQ analyzed the data. All authors contributed to the writing and editing of the paper.

**Declaration of interests:** None

**Correspondence:** juan.rocha@su.se | +46730747766

\pagebreak

# Highlights

- Historical data reveals that no country in the world has reduced inequality and environmental footprint simultanously
- A trilemma exist where countries need to trade-off progress on social and environmental goals
- An inequality trap seems to exist particularly acute in Latinamerica
- Countries trajectories on the trilemma space suggest that in the future social progress can be achieved without compromising the biosphere

\pagebreak
# Acknowledgements

We would like to thank the conveners and other participants of the 2016-2018 Beijer Young Scholars workshops for their insightful comments and advice as well as the Beijer Institute of Ecological Economics, at the Royal Swedish Academy of Sciences, for its institutional support. This work was also supported by funding from Formas (grant number 2018-02318), and the Swedish Research Council Dnr 2018-05862 and Dnr 2018-06139.

\pagebreak

# Global tradeoffs between prosperity, inequality, and the environment

\begin{abstract}

  Sustainability has often been defined as jointly achieving prosperity, equality, and environmental integrity: in other words, material living standards that are high, broadly-distributed, and low-impact. But what if there is a “sustainability trilemma” that inhibits the simultaneous achievement of these three developmental goals? We analyzed synergies and tradeoffs between these three key dimensions of sustainability using data on per-capita gross national income, the Gini coefficient for income distribution, and per-capita ecological footprint from 1995 to 2017. Over these two decades, we found that no country in the world overcame the hypothesized trilemma. Nonetheless,the trajectories of  countries in the three-dimensional space formed by these key indicators revealed  typologies that identified  common development barriers as well as potential strategies to overcome them. Historically, countries tended to decrease in income inequality before growing in prosperity and environmental impact. Some countries were trapped in a space of high inequality, while prosperous countries lowered inequality only at the expense of increasing environmental impact. Achieving  greater synergy at the nexus of these three sustainability dimensions will likely require countries - especially those that have already become prosperous - to prioritize economic redistribution and environmental stewardship alongside or even before material growth. At the global level, structural inequities in the world economy as well as historical and geographical factors suggest the need for more urgent collective action, especially in the face of the Anthropocene’s worsening risks. Without such concerted efforts, the sustainable development of individual countries, and of the world, may continue to be stymied by this trilemma.

\end{abstract}

# Introduction

Sustainability has often been characterised as having three essential and nested dimensions: economy, society, and the environment. More specifically, this tripartite framework  has been defined as development that attains high material living standards (prosperity) that are well distributed (equality) and maintainable over the long run (environmental integrity) [@Ehrlich:2012; @Sachs:2015; @Tracey:2008]. Although each goal is desirable in its own right, the question nonetheless needs to be asked: What if all three are not simultaneously achievable? It has often been assumed that the positive ends to which individuals as well as societies aspire can cohere, but *a priori*, there is no reason to believe that diverse goals must be compatible (across scales) or scalable (without upper bounds) [@Gray:2007; @Greenspan:2001].

In the literature on development, such macro-level incompatibilities are not novel. For instance, it has been argued that a country cannot achieve globalization, democracy, and sovereignty in a non-volatile combination [@RodrikD:2011]. Using resources efficiently, fairly and with maximum freedom has also been proposed as an unattainable triad [@Coyle:2011]. Additionally, based on international trends, research suggests that there may be persisting tradeoffs between dietary improvement, disease mitigation, and environmental integrity [@Clark:2018]. In this light, we argue that there may be another such "trilemma" facing countries in the Anthropocene: the difficulty - or perhaps impossibility - of being prosperous and equal while achieving or maintaining low environmental impact.

For most of human history, material want, not sufficiency, was the norm for most people regardless of county [@Maddison:2007; @Harari:2014]. Although global prosperity has increased in recent generations, progress has been uneven. If China’s gains in poverty alleviation were removed, the United Nations’ Millennium Development Goal of halving extreme poverty between 1990 and 2015 would not have been achieved [@ki2013millennium]. And even today, a significant share of the world’s population still lives under the international poverty line (a standard that the UN itself recognizes may grossly undercount the world's impoverished[@Alston_20220]). Right above this derisory threshold,  hundreds of millions more people maintain a precarious subsistence and understandably aspires for higher and more secure living standards [@TheWor:2018]. Furthermore, the recent COVID-19 pandemic pushed tens-of-millions of people back beneath that already low poverty line, leading to an increase rise of global poverty unprecedented since the end of World War II and underscoring the economic precarity that still afflicts much of the world [@buheji2020extent].

At the same time, while average incomes for certain countries in what was once called the "Third World" began to  converge with those of the advanced economies in recent decades, asymmetries in the domestic distribution of income and wealth have worsened within both developed and developing countries [@Bourguignon:2007; @Milanovic:2016].
The widespread feelings of injustice that the increasingly unequal economic distribution arouses has historically been called the "Social Question", inspiring influential theories of political transformation in the 19th and 20th Centuries (for instance various strands of Marxist and Third-Worldist thought) [@Christiansen:2019; @Faist:2009]. Over the past decade, inequality has once again become an increasingly urgent issue for policymakers as well as the general public, prompting anxieties about social and political instability [@Chancel:2020; @UN2020]. These anxieties have been reinforced by a growing body of research demonstrating complex linkages between inequality and a range of financial, health, and ecological risks[@Hamann:2018; @Leach:2018; @Pikett:2014; @Wilkinson:2009; @Hartmann:2017]. Furthermore, the impacts and policy responses from the COVID-19 pandemic worsened income and wealth inequalities within many countries, as well as widening the development trajectories between high- and low-income economies [@WorldBank:2022].

Environmental integrity forms the third, and arguably most fundamental, dimension of the sustainability trilemma [@arrow1995economic]. Interdisciplinary research in sustainability science confirms that human well-being and economic development depend on the resilience and abundance of nature. At the same time, there is now a scientific consensus that human activities have a wide and deleterious impact on the biosphere, which constitutes the habitative upper bound and common denominator for our species (notwithstanding fanciful visions of extraterrestrial expansion)  [@Díaz:2019; @folke2021our]. Anthropogenic drivers  have become primarily responsible for global biodiversity loss, land degradation, and climate change, among other environmental impacts [@Díaz:2019]. Indeed, humanity  has already exceeded a number of the biophysical boundaries that delimit a “safe operating space” for humanity [@Rockström:2009; @RaworthKate:2017]. Global development in the 21st Century therefore urgently requires more sensible stewardship of the biosphere, from ecosystems to the climate [@Millenniu:2005; @IPCC:2021].

The United Nations’ Sustainable Development Goals explicitly recognize the important nexus formed by prosperity, equality, and the environment  [@Sachs:2015]. And in its report charting the path towards a sustainable recovery in the wake of COVID-19, the World Bank also identified the need to achieve “shared prosperity” by reducing inequality and accounting for the climate crisis and other environmental challenges [@WorldBank:2022]. Nonetheless, while all three of these development goals are seen as necessary, it remains for empirical investigation to determine whether the pursuit of any one helps or hinders the pursuit of the other two. In this study, we ask whether there is evidence – and rationale – that all three goals can be jointly achieved, by any given country and by the world as a whole. In other words, this study attempts to answer the questions: To what extent does the sustainability trilemma exist and are there inherent tradeoffs between these goals? If a sustainability trilemma does exist, where does it occur and why does it occur there?

# Methods

To answer these research questions, we examined the performance of countries over the past quarter-century using publicly available data on the three dimensions of the sustainability trilemma described above (Table \ref{tab1}). We chose the country level based on the fact that many of the most consequential policies, plans, and institutions affecting sustainability are made by national governments [@Bryan:2018], and due to practical considerations of data availability. We operationalized the sustainability trilemma as the movement of countries over time in the three-dimensional space formed by the axes of prosperity, (in)equality, and environmental integrity (alternatively, impact)  (Fig \ref{fig:trilemma}). There is a development synergy if a country moves in desired directions towards more than one goal (e.g., reducing inequality and increasing prosperity). There is a development tradeoff if positive movement towards one goal is accompanied by movement in an undesirable direction for another. Therefore, synergies between two goals are movements along the (blue) diagonal (positively correlated), while tradeoffs are represented by the (red) axis, where the achievement of one hinders the achievement of another (negatively correlated). It is important to note that in this framing, the sign of the correlation changes depending on the term used. For instance, if one speaks of equality as opposed to inequality, or environmental integrity as opposed to environmental impact. In the remainder of this paper, we use inequality and environmental impact (proxied by ecological footprint) because this is how the original data was defined and structured. Therefore, across these two domains, increasing numbers mean less desirable conditions, and negative correlations translate to synergies between the goals of prosperity, equality, and environmental integrity  (Fig \ref{fig:trilemma}).

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 4in]{fig1_trilemma_space}
\caption{\textbf{Trilemma space} Countries around the world face the challenge of reducing inequality and environmental impact while increasing prosperity (A). There could be synergies and trade-offs in achieving these three objectives (B-D). We explore trajectory typologies in the trilemma space (E-G). 53 trajectories from 1995 to 2017 are depicted. The trilemma is projected in each of the facets of the tridimensional cube: inequality and ecological footprint (E), inequality and prosperity (F), and ecological footprint and prosperity (G). Thresholds that delimit the desirable space are depicted as dashed lines and the colours represent negative synergies (red), positive synergies (blue), and tradeoffs (light orange) between the dimensions of the trilemma (see Methods).}
\label{fig:trilemma}
\end{figure*}
```

To quantify prosperity, inequality and environmental impact, we identified a set of widely-used and publicly-available indicators (Table \ref{tab1}). For prosperity, we used per-capita gross national income (GNI) in US\$ values. GNI is related to gross domestic product (GDP) and the two are close in value for most countries. A key difference is that GNI accounts for earnings from foreign investments. The World Bank uses GNI per capita as its standard measure for the level of development (e.g., low-, medium- or high-income, data source: <http://data.worldbank.org>). The threshold for being "prosperous" was US\$12,746 per capita per year, which was the World Bank's definition for a high-income country in 2017, the concluding year of our study period [@TheWor:2018].

To measure inequality, we used the Gini coefficient for income. The Gini coefficient is a statistical measure of dispersion and has a long history of being used to measure the level of economic inequality within a country [@Pyatt:1976]. Hypothetically, a country with a Gini value of 0 is absolutely equal, while a value of 1 represents absolute inequality. Gini coefficients of income were taken from the World Income Inequality Database v. 3.4 (source: [https://www.wider.unu.edu/data](https://www.wider.unu.edu/data)). We used 0.30 as the threshold below which countries are identified as having low inequality, similar to findings from previous historical surveys [@Piketty:2014].

We used per-capita ecological footprint to measure environmental impact. Ecological footprint is a holistic estimate of the biophysical resources used by human activity, quantified in terms of the land area required for the extraction of materials and energy and the absorption by the environment of pollution and waste [@Wackernagel:1998]. Its unit is the "global hectare", combining measurements of carbon emissions, built-up land, fishing grounds, croplands, forests, and grazing land into an index of environmental impact. In conjunction with measurements of biocapacity – or the extent to which the environment can provide, absorb, and regenerate – the ecological footprint determines whether a given country (or person or city or the world as a whole) is in an unsustainable condition of ecological "deficit". Though it is not exhaustive, the ecological footprint provides the most widely-used and data-rich approximation of overall human impact on the environment, and it accounts for the effects of trade on resource consumption by country. Ecological footprint and biocapacity data were taken from the Global Footprint Network (source: [https://www.footprintnetwork.org/resources/data/](https://www.footprintnetwork.org/resources/data/)). We used 1.68 global hectares per capita as the threshold for environmental impact; any value higher would exceed global biocapacity (i.e., ecological deficit). In other words, it is the limit at which current generations' needs can be satisfied without compromising the ability of future generations to satisfy their needs.

```{r tab1, echo = FALSE, include = TRUE, results = "asis"}
dat <- tibble(
    Dimension = c("Prosperity", "Equality", "Environment"),
    Indicator = c(
        "Per-capita gross national income in US$",
        "Gini coefficient for income",
        "Per-capita ecological footprint in global hectares"),
    N = c(232, 196, 211),
    Threshold = c(12746, 0.3, 1.68)
)

stargazer::stargazer(
    dat, type = "latex", 
    title = "Indicators used in the assessment of national prosperity, equality and environmental impact.",
    label = "tab1", summary = FALSE, header = FALSE)

# knitr::kable(dat, caption = "Indicators used in the assessment of national prosperity, equality and environmental impact.", label = "tab1")
```

We used these data to chart the movement of countries across the dimensions of prosperity, inequality, and environmental impact over time. We further investigated whether these movements trace out typologies of countries within the trilemma space. To identify different national profiles, we calculated the magnitudes and angles of displacement of countries over time.  We then used average and standard deviation statistics to discover country clusters and performed sensitivity analyses to determine the relevant number of clusters (clustering is a statistical process whereby objects are partitioned and then grouped in such a way that the members of one cluster are more closely related to each other than they are to those of other clusters). We compared ten clustering algorithms over 30 different metrics of performance, following protocols outlined by refs [@Charrad:2014tp; @Brock:2008vz]. We used principal component analysis as an ordination method for visualization, and then vector-fitting to show how clusters are pulled apart in the principal component space (in order to make the three-dimensional distribution of countries intelligible). All data used are publicly available, and the code used for the analysis is available at: <link will be added upon publication, avoided here due to double-blind requirements>.

An analysis of the missing values in the publicly available data shows that assessing changes over time in the trilemma space was only possible for 53 countries. The most incomplete dataset is that for inequality (Figs \ref{fig:SM_gni_missing}, \ref{fig:SM_inq_missing}, \ref{fig:SM_ecofp_missing}), where for any given year at least half of the countries of the world have missing values. Based on the quality of the three datasets, we selected time series from 1995-2017 for the 53 countries with $\leqslant 30\%$ of missing values and then imputed them with a linear interpolation. This reduced dataset, although incomplete, enables the study of the dynamics of countries in the trilemma space (Fig \ref{fig:trilemma}).

# Results

No country simultaneously achieved high prosperity, low inequality, and low environmental impact. This strongly suggests that globally, there is a sustainability trilemma. At the country level, there were strong tradeoffs between reducing inequality and reducing environmental impact (Fig \ref{fig:trilemma}E - as the Gini coefficient for income decreases, ecological footprint increases) as well as between increasing prosperity and reducing environmental impact (Fig \ref{fig:trilemma}G- per-capita GNI and ecological footprint move in the same direction). Results show synergies between increasing prosperity and reducing inequality (Fig \ref{fig:trilemma}G), although a number of high-income countries (e.g., Sweden, New Zealand) exhibited low levels of inequality in the past, but also showed high variation and a trend towards greater income inequality in more recent years.

While these globally-aggregated patterns provide insight into international variations with respect to key dimensions of sustainability, they do not necessarily reveal the diversity of individual country trajectories, nor their typologies. To further investigate individual trajectories, we calculated the displacement within the trilemma space for each pair-wise dimension, where $\gamma$ (gamma) is the angle in the inequality and prosperity plane while $\theta$ (theta) is the angle in the inequality and ecological footprint plane. These two angles and the total distance covered in the three-dimensional space help to identify typologies in country trajectories.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 4in]{trilema_ensamble}
\caption{\textbf{Countries' positions and trajectories within the trilemma space} 53 country trajectories from 1995 to 2017 are depicted. Colours and contours represent areas of similar values and enclose attractors within peaks. The insets of each graph show the direction of movement for countries during the study period.}
\label{fig:cube_ensamble}
\end{figure*}
```

There are certain regions of the trilemma space where countries tended to dwell (Fig \ref{fig:cube_ensamble}). The inequality-ecological footprint plane (Fig \ref{fig:cube_ensamble}A), and prosperity-ecological footprint plane (Fig \ref{fig:cube_ensamble}C) have two of these attractors, while the inequality-prosperity plane (Fig \ref{fig:cube_ensamble}B) has three, suggesting the existence of two to three country trajectory clusters. Despite this clustering, countries showed distinctive profiles in terms of their direction of movement (Fig \ref{fig:cube_ensamble} insets). Most countries had low displacement (short arrows), meaning that their trajectory over the study period did not take them far from the region in the trilemma space where they started. A few countries were distinctive in their direction and magnitude of displacement from their starting point in 1995. For example, between 1995 and 2017, China's Gini and ecological footprint worsened while per-capita GNI grew significantly. And while inequality and environmental impact actually fell in Ukraine, prosperity only increased marginally. To better understand country typologies, we then clustered summary statistics (mean and standard deviations) of the three trilemma dimensions, as well as the direction and magnitude of country movement over time (Fig \ref{fig:pca}).

The trajectories of countries in the trilemma space from 1995 to 2017 converged into three general types (Fig \ref{fig:pca}A). First, there was a group of countries "trapped" in a region of high inequality, mainly in Latin America. They moved toward regions of higher income inequality with respect to the sample and with a larger standard deviation of $\theta$; that is to say, large variations on the inequality and ecological footprint axes (Fig \ref{fig:pca}A, in green). The second group showed large displacements within the trilemma space but with no definite direction, and was characterized by particularly high variations in its Gini coefficient for income and per-capita GNI (Fig \ref{fig:pca}A, in yellow). The third group was composed of countries with relatively higher prosperity and ecological footprint; their movements in the trilemma space were relatively well-defined, with larger values but lower variation in direction ($\gamma$ and $\theta$ angles, Fig \ref{fig:pca}A, blue group). If we ignore temporal dynamics and focus on the mean and standard deviation of the trilemma dimensions as summary statistics, we are able to include more countries (N=140). Consequently, this yielded two distinct clusters of countries (Fig \ref{fig:pca}B), with similar tradeoffs between prosperity, inequality, and environmental impact. The first cluster (Fig \ref{fig:pca}B, in yellow) was composed primarily of advanced Western economies, while the second (Fig \ref{fig:pca}B, in green) was composed of countries from Africa, Latin America, and Asia.

```{=tex}
\begin{figure*}[htp]
\centering
\includegraphics[width = 6.5in, height = 5in]{pca_maps}
\caption{\textbf{Country typologies on the trilemma space} Principal component analysis and a senstivity analysis on clustering routines are used to identify country typologies. Three clusters are identified when looking at the historical trajectory of countries (N=53) from 1995-2017 (A). Two clusters are identified when aggregating over time with means and standard deviations on the trilemma dimensions regardless of data completeness. It complements the dynamic analysis with one summary statistic over time for 140 countries (B). For both A and B, red arrows are vectors fitted to the data that indicate the directions along which countries and clusters differ from each other. Mean values of ecological footprint (C), inequality (D), and gross national income (E) are shown in maps for the 140 countries.}
\label{fig:pca}
\end{figure*}
```

# Discussion

Overall, our analysis found that decreasing inequality and decreasing environmental impact pulled countries in opposite directions (Fig \ref{fig:pca}), providing strong evidence for the difficulty of achieving both goals at the same time. Also, countries typically decreased in inequality before increasing in environmental impact (Fig \ref{fig:cube_ensamble}A). Similarly, countries tended to see reductions in inequality before prosperity rose (Fig \ref{fig:cube_ensamble}B). Importantly, many middle-income countries had levels of inequality similar to high-income countries, but at lower levels of environmental impact and prosperity (Fig \ref{fig:cube_ensamble}B, C). Overall, our findings suggest that higher prosperity usually comes at the expense of environmental integrity. National trajectories through the three-dimensional trilemma space  also revealed that some countries struggled to escape areas characterized by high inequality, which may represent an “inequality trap” subset of the larger sustainability trilemma.

It is important to note that this study was a historical, not predictive, analysis, and that past trajectories do not necessarily determine plausible future pathways. The potential mechanisms behind these tradeoffs and synergies between prosperity, inequality, and environmental impact constitute an important research frontier in sustainability science and development studies. The patterns and typologies identified in this paper could provide insights or directions for future investigations. Nonetheless, some mechanisms have already been hypothesized for pairwise relationships among the indicators in the often disparate literatures of economics, sociology, and environmental science. In what follows, we review several of the most influential hypotheses in the context of our findings.

Tradeoffs between inequality and prosperity (e.g., per-capita income) have been extensively studied in relation to the concept of the Kuznets Curve (KC). In this framing, the former increases as the latter grows, eventually reaching an inflection point after which growing prosperity leads to lower inequality [@Kuznets:1955]. The evidence for this hypothesis has been conflicting, with some country-level studies finding strong correlations and others detecting no clear relationships [@Chancel:2020]. This has made the identification of mechanisms difficult. Indeed, sustained inequality reduction over the past century seems to have been a consequence of contingent but systemic shocks - the world wars and the Great Depression - rather than through regularities of the global economy or concerted government actions  [@Pikett:2014]. Our analysis found no evidence for a KC-like relationship in our study period, which was a time of rapid economic growth in many parts of the world.

In terms of the economy-environment relationship, the Environmental Kuznets Curve (EKC) has often been advanced as an axiomatic relationship: countries increase in environmental impact with growing prosperity, but this trend reverses after reaching an inflection point of affluence[@barbier1997economic]. As with the original KC, empirical studies of a wide range of variables have come to conflicting conclusions. While some identified international evidence for this inverted-U-shaped relationship, others did not [@dasgupta2002confronting; @stern2004rise; @pothen:2019]. In the trilemma space, with the composite indicator of ecological footprint as the environmental indicator, we also did not detect an EKC-type trend: i.e., an inflection point in the prosperity-environmental impact plane (Fig \ref{fig:cube_ensamble}). Various mechanisms have been proposed to explain this relationship (or its absence), including market liberalization, more stringent regulations, and greater freedoms for citizens to voice their concerns [@MALER201325]. But what is clear from diverse empirical studies is that the connections between prosperity and environmental impact are mediated or confounded by other, largely social and institutional, factors.

Notably, structural inequities in the global economy are creating or exacerbating a skewed distribution of development burdens. Environmental impacts can be "outsourced", as is common with affluent countries that have shifted resource-intensive production to developing countries through global supply chains [@Lambin:2011] - an important factor not captured in most EKC analyses [@MALER201325]. For instance, the US and a number of European countries have outsourced a significant portion of their carbon emissions, land use, and environmental impacts to lower-income economies through trade [@Nijdam:2005; @Peters:2006; @Weinzett:2013]. Additionally, it has been found that international trade significantly influences global land use patterns, worsening natural resource exploitation and biodiversity loss [@Creutzig:2019].

Within the trilemma space, there were international clusters that moved in distinctively different directions. These divergences further highlight the inadequacy of reductionistic, especially mono-causal, explanations (such as those found in many KC and EKC studies) and the need to incorporate historical and geographic factors as well. For instance, at the end of the study period, Ukraine had (relatively) low levels of inequality and environmental impact. However, the country’s  trajectory through the trilemma space was turbulent: over the study period, indicators rose and fell in dramatic downswings due to market shocks and corruption, endemic across many post-Soviet countries [@Duvanova:2014], as well as the social and economic dislocations caused by violent conflict. It is noteworthy that it was only in 2013  when Ukraine’s ecological footprint fell below its biocapacity (Fig. \ref{fig:S4}). Countries that performed relatively well across all three dimensions tended to be advanced Western economies, particularly those along the North Atlantic rim (e.g., Canada, Sweden). These countries had a much higher ratio of space and resources to people, as reflected in per-capita biocapacity.

Interactions between the structure of the global economy, geographic conditions, and historical contingencies can help explain international differences with respect to the sustainability trilemma. This is perhaps best illustrated by the country that underwent the greatest magnitude of change in all three dimensions during the study period. In China, the rapid rise of prosperity, inequality, and environmental impact were significantly driven by the country’s integration into international commercial networks (Fig. \ref{fig:S4}). The rapid growth of incomes was underpinned by a resource-intensive export sector while rising income inequality was due in large part to geographic segmentation: urban coastal areas were the affluent export hubs while more rural inland areas supplied raw materials, intermediate goods, and cheap labour [@xie2014income; @qi2013interpreting]. These changes, and the subsequent surges in all three dimensions, were set in motion after the tumultuous historical events of the 1960s and 1970s, which prompted the national leadership to undertake extensive economic reforms.

Additionally, complex interactions among the three dimensions of the trilemma - and their implications for development - have also been recognized in interdisciplinary studies (none of which posit simple linear or quadratic relationships between the economic, social, and environmental dimensions of sustainability). For instance, Weinzettel et al. [@Weinzett:2013] found that countries with high per-capita biocapacity (i.e., large natural resource endowments relative to population level) tended to spare more land for nature, forming a virtuous cycle of improving environmental conditions. Similarly, population, environmental carrying capacity, and pollution exhibited mutually-reinforcing feedbacks across countries [@LaTorr:2019]. These types of feedbacks can create and maintain "red loop" or "green loop" dynamics between poverty and the environment, entrapping some countries in long-term conditions of high poverty and environmental degradation while placing others on a more sustainable trajectory [@Cumming:2014; @Cumming:2018; @Hamann:2015].

Although our analysis identified important patterns and drivers of international development (and corresponding lacunae in discipline-specific literature), it still has several limitations. First, the offshoring of environmental degradation discussed above is not completely captured in standard ecological footprint measurements [@VandenBergh:2012]. A consumption-based approach to measuring environmental impacts – as opposed to the more mainstream production-based approach – may help internalize at least some of these externalities (e.g., many resource-intensive goods produced in China and Vietnam are consumed in the US and Europe, but the increases in ecological footprint thus generated are attributed to the former two). However, such an approach is constrained by the lack of relevant data, with  existing literature tending to be more narrowly focused on specific commodities and on bilateral or regional trade. If these outsourced impacts were incorporated into the appropriate indicators, our findings would likely show that affluent countries have been even less successful in overcoming the sustainability trilemma. Addressing this issue has important implications for the international distribution of responsibilities with respect to global commons challenges such as  mitigating climate change, biodiversity loss, and pandemic risks - all of which disproportionately affect lower-income countries [@Fankhauser:2014; @Mendelsohn:2006; @IPCC:2021].

We also recognize that there are other valid ways to conceptualize sustainability. For instance, studies on the relationship between ecological limits, human population, and the economy have a long pedigree [@Ehrlich:1971; @Meadows:1972]. More recently, a popular  framework put forward by Raworth [-@RaworthKate:2017] postulated 11 "social thresholds" and 7 planetary boundaries to delimit sustainability. This framework and similar approaches imply equal weight among chosen thresholds, which mix subjective indicators (e.g., life satisfaction and democratic quality) with objective measures of wellbeing (e.g., nutrition and life expectancy). Such classification schemes are informative but may also be overdetermined: it is conceivable that a different and even longer list of thresholds could be advanced to form an equally relevant sustainability framework. In addition, there have been calls for analyses to better integrate social and ecological goals in order to account for the foundational role played by nature [@reyers2020global]. While acknowledging that our “classic” three-pillars framework (sometimes termed the “triple bottom line”) is imperfect, we believe it nonetheless captures long-standing and important dimensions of sustainability and how countries have developed with respect to them.

# Conclusions

Our study found that countries around the world face a trilemma when it comes to achieving economic prosperity and social equality while reducing environmental impact - three goals that development scholarship and international agendas have defined as being fundamental to sustainability. It is important to recognize how institutional constraints, geographic differences, and historical contingencies and their interactions vary across countries. When it comes to such complex, multi-dimensional challenges, there are no "one-size-fits-all" solutions [@Ostrom:2007; @ostrom2010moving; @perrings2007future]. We therefore argue that national approaches to overcoming the sustainability trilemma will need to be context-specific and must account for path-dependence. At the same time, our findings indicate that there are typologies of how countries have historically evolved with respect to prosperity, inequality, and the environment. These typologies can help identify common priorities and development strategies. This may be particularly relevant for Global South countries, which appear to face the steepest challenges with respect to the sustainability trilemma.

Whether the synergies and tradeoffs between prosperity, inequality, and environmental impact identified in this study hold across scales (e.g., at finer subnational scales or in other analytic units such as cities or corporations), remains an open question. Furthermore, these mechanisms  are likely to be complex rather than mono-causal [@Weinzett:2013; @Cumming:2018; @Hamann:2018; @LaTorr:2019]. Additionally, the foundational role of nature, which provides the basic life-support systems for humanity, cannot be understated when thinking about sustainable development. As our findings strongly suggest, a myopic focus on ever-growing prosperity as the solution to inequality and environmental degradation is misplaced - neither KC or EKC relationships were found. Indeed, highly prosperous countries do not necessarily have lower levels of inequality, but almost always have higher environmental impacts. One undeniable trend from recent history is that a singular focus on expanding income  can undermine progress towards the ambition of leaving no country or person behind in the pursuit of global sustainability [@arrow2004we; @fanning2022social].

Overcoming the sustainability trilemma, for individual countries and the world as a whole, demands development that focuses as much on the distribution and impact of the economy as on its continued growth. This will likely require much greater investment in the social and natural capital that have been degraded, or allowed to atrophy, across the world in recent decades [@pretty2003social; @brondizio2009connectivity; @wackernagel1997perceptual; @Ehrlich:2012]. Through this more inclusive approach, countries could still become more prosperous - just in a way that distributes those gains more fairly, which would help resolve the poverty and precarity that still bedevil so many, and with less impact on the biosphere, which is the ultimate guarantor of humanity’s prospective flourishing.

# References {.unnumbered}

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{prosperity_missing_values}
\caption{\textbf{Data quality for prosperity} Different prosperity metrics are compared with respect to their coverage in space (countries) and time (A). Gross national income (GNI) is the metric with higher completion rate (B), with an arbitrary threshold of 30\% missing values, the analysis could include over 100 coutries (C).}
\label{fig:SM_gni_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{inq_missing_values}
\caption{\textbf{Data quality for inequality} Mean Gini income is the most complete statistic of inequality with respect to their coverage in space (countries) and time (A). Alternative metrics such as the Palma index overcome some of the limitations of Gini but have disproportionally more missing values (B). For Gini, the proportion of countries with missing values is above 50\% for any given year. An arbitrary treshold of 30\% threshold for includsion will leave countries like South Africa (country with highers inequality in the world) out of the analysis (D).}
\label{fig:SM_inq_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{ecofp_missing_values}
\caption{\textbf{Data quality for ecological footprint} Coverage of the ecological footprint in space (countries) and time (A) shows that data is relatively complete after mid 1990s (B) An arbitrary threshold of 10 missing values would leave some of the small islands states out of the analysis (C).}
\label{fig:SM_ecofp_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{FigS4}
\caption{\textbf{Trajectories examples} As shown by the experiences of Ukraine and China from 1995 to 2017, historical and structural factors influence the trajectories of countries within the "trilemma space" formed by the economic, social, and environmental dimensions of sustainability: (A) the global financial crisis; (B) the beginning of the Russo-Ukrainian conflict; (C) the accession of China into the World Trade Organization; (D) the beginning of the large-scale anti-corruption crackdown in China.}
\label{fig:S4}
\end{figure*}
```