### 2023-11-20 revised_manuscript_231115.Rmd
uploading replication datasets

---
title: "Triple bottom line or trilemma? Global tradeoffs between prosperity, inequality, and the environment"
author: |
    | \small Tong Wu^1,§^, Juan C. Rocha^2,3,4§^, Kevin Berry^5^, Tomas Chaigneau^6^,  Maike Hamann^7,8^, Emilie Lindkvist^2^, Jiangxiao Qiu^9^, 
    | \small Caroline Schill^2,4,10^, Alon Shepon^11^, Anne-Sophie Crépin^10^, and Carl Folke^2,4,10^
    | \small
    | \footnotesize ^1^The Natural Capital Project, Doerr School of Sustainability, Stanford University
    | \footnotesize ^2^Stockholm Resilience Centre, Stockholm University
    | \footnotesize ^3^Future Earth
    | \footnotesize ^4^ The Anthropocene Laboratory, The Royal Swedish Academy of Sciences
    | \footnotesize ^5^College of Business and Public Policy, University of Alaska Anchorage
    | \footnotesize ^6^Environment and Sustainability Institute, University of Exeter
    | \footnotesize ^7^Centre for Sustainability Transitions, Stellenbosch University
    | \footnotesize ^8^Centre for Geography and Environmental Science, University of Exeter
    | \footnotesize ^9^School of Forest Resources & Conservation, Fort Lauderdale Research and Education Center, University of Florida
    | \footnotesize ^10^Beijer Institute of Ecological Economics, The Royal Swedish Academy of Sciences
    | \footnotesize ^11^Department of Environmental Studies, The Porter School of the Environment and Earth Science, Tel Aviv University
    | \footnotesize ^§^These authors made equal contributions
#   | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/apa.csl
bibliography: paper/references_ei.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

```{r data, include=FALSE}
library(tidyverse)
library(here)

## word counts
## Abstract 150
## Intro 697
## Method 957
## Results 621
## Discussion 1434
## Conclusion 267
## Total main 3976 + 150 abstract
## 
## Targets: 
# 
# - NatComms: 5000w | 150w abstract | 4 figs or tables | Methods end 3000w | Refs
# 70max
# 
# - WorldDev: | 4500-10k words including all (refs, captions, tables, etc) | 300w
# abstract
```

\doublespacing

**Author contributions:**  Conceptualization was a team effort, TW and JCR produced the first manuscript, TW, JCR and JQ analyzed the data. All authors contributed to the writing and editing of the paper.

**Declaration of interests:** None

**Correspondence:** juan.rocha@su.se | +46730747766

\pagebreak

# Highlights

- No country in the world simultaneously achieved high prosperity, low inequality, and low environmental impact between 1995-2017.
- A sustainability “trilemma” as opposed a “triple bottom line” appears to exist for international development.
- There is an “inequality trap” that appears to be particularly acute in Latin America.
- Country trajectories suggest that reducing inequality can be achieved without compromising the biosphere.
- The three dimensions need be considered together to identify international development challenges.

\pagebreak
# Acknowledgements

We would like to thank the conveners and other participants of the 2016-2018 Beijer Young Scholars workshops for their insightful comments and advice as well as the Beijer Institute of Ecological Economics, at the Royal Swedish Academy of Sciences, for its institutional support. This work was also supported by funding from Formas (grant number 2018-02318 and grant number 2020-00454), and the Swedish Research Council Dnr 2018-05862 and Dnr 2018-06139.

\pagebreak

# Triple bottom line or trilemma? Global tradeoffs between prosperity, inequality, and the environment

\begin{abstract}

A key aim of sustainable development is the joint achievement of prosperity, equality, and environmental integrity: in other words, material living standards that are high, broadly-distributed, and low-impact. This has often been called the “triple bottom line”. But instead, what if there is a " trilemma" that inhibits the simultaneous achievement of these three goals? We analyzed international patterns and trends in the relationships between per-capita gross national income, the Gini coefficient for income distribution, and per-capita ecological footprint from 1995 to 2017. Based on these widely-used measures of development, we found that no country achieved the triple bottom line. Over the study period, countries tended to decrease in income inequality before growing in prosperity and environmental impact. A cluster of countries, primarily from Latin America, appeared trapped in a regime of high inequality, while the instances of prosperous countries lowering inequality only coincided with increasing environmental impact. The existence of country clusters and widely-divergent national trajectories show that common pair-wise explanations such as Kuznets Curves do not fully capture important development dynamics, especially in a period of global environmental change. Instead, frameworks need to consider all three dimensions together, and especially the relationship between inequality and the environment. Additionally, the past development trajectories of countries, how they fit into the structure of the world economy, and the role of social-ecological feedbacks in strengthening clustering effects such as “inequality traps”. Practically, our findings suggest that to overcome the sustainability trilemma, many countries - especially those that are already prosperous - should prioritise social redistribution and environmental stewardship before further economic growth.

\end{abstract}

# Introduction

Sustainability has often been characterised as having three essential and nested dimensions: economy, society, and the environment - what is sometimes referred to as the "triple bottom line". More specifically, an important development objective within this framework is to achieve high material living standards (prosperity) that are also well distributed (equality) and maintainable over the long run (environmental integrity) [@Ehrlich:2012; @Sachs:2015; @Tracey:2008]. Programmatically, the United Nations’ Sustainable Development Goals explicitly recognize the important nexus formed by prosperity, equality, and the environment [@Sachs:2015]. And in its report charting the path towards a sustainable recovery in the aftermath of COVID-19, the World Bank also identified the need to achieve "shared prosperity" by reducing inequality and addressing environmental crises [@WorldBank:2022]. However, whether the three goals can be simultaneously achieved remains an open question – what if there is a "sustainability  trilemma" that prevents countries from reaching this development objective?

Although global prosperity has increased in recent generations, progress has been uneven. Even today, a significant share of the world’s population lives under the international poverty line (a standard that the UN itself recognizes may grossly undercount the world's impoverished [@Alston_20220]). Right above this derisory threshold, hundreds of millions more people maintain a precarious subsistence and aspire for higher and more secure living standards [@TheWor:2018]. And while average incomes for certain developing countries have been converging with those of advanced economies in recent decades, asymmetries in the domestic distribution of income and wealth have worsened across the world [@Bourguignon:2007; @Milanovic:2016]. This has made  inequality an urgent issue for policymakers and the general public due to  anxieties about social and political instability [@Chancel:2020; @UN2020]. Environmental integrity forms the third, and perhaps most fundamental, dimension of the sustainability [@arrow1995economic]. There is now a scientific consensus that human activities have had severely deleterious impacts on the biosphere, the health of which determines long-run human wellbeing [@Díaz:2019; @folke2021our]. Indeed, humanity has already exceeded a number of the biophysical boundaries that delimit a "safe operating space" for humanity [@Rockström:2009; @RaworthKate:2017; @Rockstr_m_2023].

Identifying global patterns and trends in the relationships between prosperity, equality, and environmental impact can provide insights into the development challenges facing individual  countries as well as shared obstacles and opportunities internationally. Using a time-series dataset of representative development indicators, this study asks whether there is evidence that prosperity, equality, and environmental integrity has been jointly achieved by any country, and then explores potential mechanisms of trade-offs and synergies between the three goals. In other words, this study attempts to answer the questions: To what extent does a "sustainability trilemma" exist, and if does exist, where does it occur and why does it occur there? By providing answers to these questions, we hope to contribute to the increasingly urgent discourse on how to achieve sustainable and inclusive prosperity in an age of unprecedented planetary change.

# Methods

To assess the extent, distribution, and evolution of the hypothesized sustainability trilemma at the global level, we examined the performance of countries over the past quarter-century using publicly available data for three representative and widely-used indicators of development (Table \ref{tab1}). We chose the country level because many of the most consequential policies, plans, and institutions affecting sustainability are made by national governments [@Bryan:2018], and also due to practical considerations of data availability. We operationalized the sustainability trilemma as the movement of countries over time in the three-dimensional space formed by the axes of prosperity, (in)equality, and environmental integrity (alternatively, impact)  (Fig \ref{fig:trilemma}). There is a development synergy if a country moves in desired directions towards more than one goal (e.g., reducing inequality and increasing prosperity). There is a development tradeoff if positive movement towards one goal is accompanied by movement in an undesirable direction for another. Therefore, synergies between two goals are movements along the (blue) diagonal (positively correlated), while tradeoffs are represented by the (red) axis, where the achievement of one hinders the achievement of another (negatively correlated). It is important to note that in this framing, the sign of the correlation changes depending on the term used. For instance, if one speaks of equality as opposed to inequality, or environmental integrity as opposed to environmental impact. In the remainder of this paper, we use inequality and environmental impact because this is how the original data was defined and structured. Therefore, across these two domains, increasing numbers mean less desirable conditions, and negative correlations translate to synergies between the goals of prosperity, equality, and environmental integrity  (Fig \ref{fig:trilemma}).

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 4in]{fig1_trilemma_space}
\caption{\textbf{Trilemma space} Countries around the world face the challenge of reducing inequality and environmental impact while increasing prosperity (A). There could be synergies and trade-offs in achieving these three objectives (B-D). We explore trajectory typologies in the trilemma space (E-G). 59 country trajectories from 1995 to 2017 are depicted. The trilemma is projected in each of the facets of the tridimensional cube: ecological footprint and prosperity (E), inequality and prosperity (F), and inequality and ecological footprint (G). Thresholds that delimit the desirable space are depicted as dashed lines and the colours represent negative synergies (red), positive synergies (blue), and tradeoffs (light orange) between the dimensions of the trilemma (see Methods).}
\label{fig:trilemma}
\end{figure*}
```

To quantify prosperity, inequality and environmental impact, we identified a set of widely-used and publicly-available indicators (Table \ref{tab1}). For prosperity, we used per-capita gross national income (GNI) in US\$ values. GNI is related to gross domestic product (GDP) and the two are close in value for most countries. A key difference is that GNI accounts for earnings from foreign investments. The World Bank uses GNI per capita as its standard measure for the level of development (e.g., low-, medium- or high-income, data source: <http://data.worldbank.org>). The threshold for being "prosperous" was US\$12,746 per capita per year, which was the World Bank's definition for a high-income country in 2017, the concluding year of our study period [@TheWor:2018].

To measure inequality, we used the Gini coefficient for income. The Gini coefficient is a statistical measure of dispersion and has a long history of being used to measure the level of economic inequality within a country [@Pyatt:1976]. Hypothetically, a country with a Gini value of 0 is absolutely equal, while a value of 1 represents absolute inequality. Gini coefficients of income were taken from the World Income Inequality Database v. 3.4 (source: [https://www.wider.unu.edu/data](https://www.wider.unu.edu/data)). We used 0.30 as the threshold below which countries are identified as having low inequality, based on assessments from historical economic studies [@Piketty:2014].

We used per-capita ecological footprint to measure environmental impact. Ecological footprint is a holistic estimate of the biophysical resources used by human activity, quantified in terms of the land area required for the extraction of materials and energy and the absorption by the environment of pollution and waste [@Wackernagel:1998]. Its unit is the "global hectare", combining measurements of carbon emissions, built-up land, fishing grounds, croplands, forests, and grazing land into an index of environmental impact. In conjunction with measurements of biocapacity – or the extent to which the environment can provide, absorb, and regenerate – the ecological footprint determines whether a given country (or person or city or the world as a whole) is in an unsustainable condition of ecological "deficit". Although it may be too aggregative of the diverse facets of anthropogenic environmental change, the ecological footprint remains the most widely-used and data-rich approximation of overall human impact on the biosphere in a single metric, while also accounting for the effects of trade on resource consumption by country. Ecological footprint and biocapacity data were taken from the Global Footprint Network (source: [https://www.footprintnetwork.org/resources/data/](https://www.footprintnetwork.org/resources/data/)). We used 1.68 global hectares per capita as the threshold for environmental impact; any value higher would exceed global biocapacity (i.e., leading to ecological deficit). In other words, it is the limit at which current generations' needs can be satisfied without compromising the ability of future generations to satisfy their needs.

```{r tab1, echo = FALSE, include = TRUE, results = "asis"}
dat <- tibble(
    Dimension = c("Prosperity", "Equality", "Environment"),
    Indicator = c(
        "Per-capita gross national income in US$",
        "Gini coefficient for income",
        "Per-capita ecological footprint in global hectares"),
    N = c(232, 196, 211),
    Threshold = c(12746, 0.3, 1.68)
)

stargazer::stargazer(
    dat, type = "latex", 
    title = "Indicators used in the assessment of national prosperity, equality and environmental impact.",
    label = "tab1", summary = FALSE, header = FALSE)

# knitr::kable(dat, caption = "Indicators used in the assessment of national prosperity, equality and environmental impact.", label = "tab1")
```

We used these data to chart the movement of countries across the dimensions of prosperity, inequality, and environmental impact over time. We further investigated whether these movements trace out typologies of country trajectories within the three-dimensional space. To identify different national profiles, we calculated the magnitudes and angles of displacement of countries over time. We then used average and standard deviation statistics to discover country clusters and performed sensitivity analyses to determine the relevant number of clusters (clustering is a statistical process whereby objects are partitioned and then grouped in such a way that the members of one cluster are more closely related to each other than they are to those of other clusters). We compared ten clustering algorithms over 30 different metrics of performance, following protocols outlined by Charrad _et al._, [-@Charrad:2014tp] and Brook _et al._ [-@Brock:2008vz]. We used principal component analysis as an ordination method for visualization, and then vector-fitting to show how clusters are pulled apart in the principal component space (in order to make the three-dimensional distribution of countries intelligible). All data used are publicly available, and the code used for the analysis is available at: <link will be added upon publication, avoided here due to double-blind requirements>.

While we were able to assess 140 countries "statically" (i.e., ignoring temporal dynamics), an analysis of the missing values showed that analyzing changes over time in the trilemma space was only possible for 59 countries. The most incomplete dataset is that for inequality (Figs \ref{fig:SM_gni_missing}, \ref{fig:SM_inq_missing}, \ref{fig:SM_ecofp_missing}), where for any given year at least half of the countries of the world have missing values. Based on the quality of the three datasets, we selected the time-series from 1995-2017 for the 59 countries with fewer than 30\% of missing values and then imputed the missing values with linear interpolation. This reduced dataset enables the study of the temporal dynamics of countries in the trilemma space (Fig \ref{fig:trilemma}).

# Results

Our analysis found that no country simultaneously achieved high prosperity, low inequality, and low environmental impact. This strongly suggests that globally, there is a sustainability trilemma. At the country level, there were widespread tradeoffs between reducing inequality and reducing environmental impact (Fig \ref{fig:trilemma}E - as the Gini coefficient for income decreases, ecological footprint increases) as well as between increasing prosperity and reducing environmental impact (Fig \ref{fig:trilemma}G - per-capita GNI and ecological footprint move in the same direction). Results show synergies between increasing prosperity and decreasing inequality (Fig \ref{fig:trilemma}G), although a number of high-income countries (e.g., Sweden and New Zealand) exhibited low levels of inequality in the past but then high variation and a trend towards greater inequality in more recent years.

While these global patterns provide insight into international variations in key dimensions of sustainability, they do not completely reveal the diversity of individual country trajectories, nor their typologies. To further investigate country trajectories, we calculated the displacement within the three-dimensional space for each pair-wise dimension, where $\gamma$ (gamma) is the angle in the inequality and prosperity plane while $\theta$ (theta) is the angle in the inequality and environmental impact plane. These two angles and the total distance "travelled" in the three-dimensional space help to define typologies in country trajectories.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 7in, height = 5in]{trilema_ensamble2}
\caption{\textbf{Countries' positions and trajectories within the trilemma space} 59 country trajectories from 1995 to 2017 are depicted. Colours and contours represent areas of similar values and enclose regimes (attractors). The bottom panel shows the direction of movement for countries during the study period for each facet of the trilemma cube. Angles and distances are calculated in the three-dimensional space, but visualised in two dimensions for readability. The arrows shows the direction and magnitude of movement if all countries would have started at the same point of origin (to enable comparison). Arrow colours on the bottom panel correspond to groups found on the clustering analysis.}
\label{fig:cube_ensamble}
\end{figure*}
```

There are certain regions in the three-dimensional space where countries tended to dwell (Fig \ref{fig:cube_ensamble}). Multiple stable states can exist if feedbacks between development drivers are present. Statistically, this translates into multi-modal distributions and so-called "regimes". The prosperity-environment plane (Fig \ref{fig:cube_ensamble}A), and inequality-environment plane (Fig \ref{fig:cube_ensamble}C) have two pronounced clusters of countries (the movement of countries during the study period stayed within a relatively circumscribed space), while the inequality-prosperity plane (Fig \ref{fig:cube_ensamble}B) has one major cluster and one looser grouping. Countries also showed distinctive profiles in direction of movement (Fig \ref{fig:cube_ensamble} insets). Most countries had low displacement (short arrows), which means their trajectory over the study period did not take them far from where they started. A few countries were distinctive in their direction and magnitude of displacement from their starting point in 1995. For example, between 1995 and 2017, China's Gini coefficient for income and per-capita ecological footprint worsened while per-capita GNI grew significantly. And while inequality and environmental impact fell in Ukraine, prosperity only increased marginally. To better understand country typologies, we clustered summary statistics (i.e., mean and standard deviations) of the three indicators, as well as the direction and magnitude of country movement over time (Fig \ref{fig:pca}).

Combining the three planes, the trajectories of countries from 1995 to 2017 can be categorized into three general types (Fig \ref{fig:pca}A). First, there was a group composed of countries with relatively high prosperity and ecological footprint; their movements were relatively less erratic and had larger values but lower variation in direction ($\gamma$ and $\theta$ angles, Fig \ref{fig:pca}A, blue group). This cluster can be called "Regime A". Next, there was a group of countries, mainly from Latin America, that seemed to be "trapped" in a region of high inequality during the study period. They moved toward regions of higher income inequality with respect to the sample and with a larger standard deviation of $\theta$; that is to say, large variations along the income inequality and ecological footprint axes (Fig \ref{fig:pca}A, in green). Within the three-dimensional space, this cluster of countries with high inequality and relatively low prosperity and environmental impact can be called "Regime B". A third group of countries showed large displacements within the trilemma space but in no definite direction, and was characterized by large variations in prosperity and inequality (Fig \ref{fig:pca}A, in yellow).

If we ignore temporal dynamics and focus on the mean and standard deviation of the indicators, we are able to include more countries (N=140). This static analysis also yielded two distinct clusters of countries (Fig \ref{fig:pca}B), with similar tradeoffs between prosperity, inequality, and environmental impact as in the findings from the dynamic analysis. The first cluster (Fig \ref{fig:pca}B, in yellow) was composed primarily of advanced Western economies and had high prosperity, inequality, and environmental impact. The second cluster (Fig \ref{fig:pca}B, in green) was composed of countries from Africa, Latin America, and Asia - broadly, the "Global South". Here, countries all tended to have lower prosperity (e.g., low- or middle-income status), but among them exhibited large variations in inequality and environmental impact.

In both the dynamic and the static analysis, decreasing inequality and decreasing environmental impact appear to pull countries in opposite directions (Fig \ref{fig:pca}), providing evidence for the difficulty of simultaneously achieving both goals. In Fig \ref{fig:pca}A, the fitted vectors (arrows) show the direction in which variables explain the variance in the data and thus countries' differences. While per-capita ecological footprint and GNI point in similar directions (countries tended to increase both at the same time), the Gini for income points in the opposite direction. This suggests that countries that reduce inequality do so after increasing in environmental impact and prosperity. In the ordination performed in the larger static sample (Fig \ref{fig:pca}B), we find a similar pattern. In fact, the trajectories analysis found that countries that reduced inequality subsequently increased in environmental impact (Fig \ref{fig:cube_ensamble}A). Reducing inequality was often followed by rising prosperity (Fig \ref{fig:cube_ensamble}B), although some middle-income countries maintained relatively low levels of inequality (at levels similar to higher-income countries) while still maintaining lower levels of environmental impact (Fig \ref{fig:cube_ensamble}B, C).

```{=tex}
\begin{figure*}[htp]
\centering
\includegraphics[width = 6.5in, height = 5in]{pca_maps}
\caption{\textbf{Country typologies on the trilemma space} Principal component analysis and a senstivity analysis on clustering routines are used to identify country typologies. Three clusters are identified when looking at the historical trajectory of countries (N=59) from 1995-2017 (A). Two clusters are identified when aggregating over time with means and standard deviations regardless of data completeness. This complements the dynamic analysis with one summary statistic over time for 140 countries (B). For both A and B, red arrows are vectors fitted to the data that indicate the directions along which countries and clusters differ from each other. Mean values of per-capita ecological footprint (C), Gini for income (D), and per-capita gross national income (E) are shown in maps for the 140 countries.}
\label{fig:pca}
\end{figure*}
```

# Discussion

Understanding the patterns and trends in the relationships between prosperity, inequality, and environmental impact is an important topic for sustainability science and development research. In the existing literatures of economic, social, and environmental science, various arguments have been advanced about the relationships between the three goals. Most commonly, mechanisms have been hypothesized for pair-wise relationships, with Kuznets Curve-type relationships being the most widely analyzed. The (Environmental) Kuznets Curve posits that growing prosperity leads to worsening inequality and environmental impact, before reaching an inflection point after which the relationship is reversed. The evidence for this hypothesis has been contested with respect to both inequality [@Chancel:2020][@Pikett:2014] and the environment [@stern2004rise; @stern1996economic ]. For the latter, conclusions have often varied based on the indicator chosen to proxy environmental impact [@MALER201325]. Our analysis did not detect KC-type relationships between prosperity and either the Gini coefficient for income or ecological footprint Fig \ref{fig:S5}) (please see SI for more details). Instead, the findings strongly suggest that growing prosperity is almost always associated with worsening inequality and environmental impact.

What is clear from earlier empirical studies of pair-wise relationships of prosperity, inequality, and the environment is that whatever tradeoffs and synergies that could be identified were mediated by exogenous factors [@dasgupta2002confronting; @pothen:2019]. For instance, in his survey of relevant studies, Chancel [-@Chancel:2020] concludes that what empirical evidence has been found to support (E)KC-type relationships are actually artefacts particular to post-World War II datasets or transient historical trends with limited applicability to today’s inequality dynamics. It is only by considering the economic, social, and environmental dimensions together, and the different ways in which countries have evolved with respect to them, can we arrive at a more comprehensive understanding. In particular, while the prosperity-inequality and prosperity-environment relationships have been widely analyzed (with mixed conclusions), the inequality-environment dynamic has been less explored – and should also be considered with respect to the potentially mediating effects of the third dimension of prosperity. In our findings, this necessity is underscored by the existence of international clusters that moved in distinctively different directions.

Overall, the low frequency with which countries were able to meet even two goals – no countries in the case of the prosperity-environment relationship – in both the dynamic and static analysis is notable. This suggests that there is indeed a sustainability trilemma facing global development. In other words, despite its prominence as a definition of sustainability, our study indicates it is highly difficult, and potentially infeasible, for countries to reach the triple bottom line of prosperity, equality, and environmental integrity. However, the burden of overcoming the trilemma is not even across countries. Another important finding of the analyses is revealed through national trajectories in the three-dimensional space. Over the study period, some countries, primarily from Latin America, struggled to escape areas characterised by high inequality, which may represent an "inequality trap" subtype of the trilemma.

The existence of the trilemma in general, and of the inequality trap in particular, could be better explained through an understanding of how inequality and environmental impact are related. A potentially useful approach is to investigate how different levels of income distribution impact environmental integrity at different levels of average income (thereby incorporating the prosperity dimension). In the existing literature on inequality-environment dynamics, two broad types of mechanisms have been posited: those that are "economic" (based on assumptions about the consumption decisions of rational individuals with different relative incomes) and those that are "political" (based on assumptions about how income distribution influences collective decision-making in the form of public policy) [@berthe2015mechanisms; @@Hamann:2018].

The "economic" (often called the "marginal propensity to emit") hypothesis argues that higher inequality lessens environmental impact because a more equal distribution raises overall consumption levels. This is based on the conventional microeconomic assumption that there is a higher marginal propensity to consume among the poor than among the rich. The political – sometimes "political economy" – hypothesis was first posited by Boyce [-@boyce1994inequality] and argues that by concentrating power in the hands of the rich, economic inequality worsens environmental conditions. One mechanism is that the well-heeled tend to have a vested interest in stopping regulations and laws that protect the environment (and thereby lower profits). This has been supported by agent-based modelling studies such as that by Kempf and Rossignol [-@KEMPF_2007]. Another hypothesized mechanism comes from the opposite side of the social spectrum, in which high inequality drives the poor to engage in more environmentally-destructive activities to meet their basic needs [@zhao2021analysis].

Some studies have found greater empirical evidence for the economic explanation than the political (economy) mechanisms. In a study of Sub-Saharan countries from 1995 to 2018, Gimba _et al._ [-@gimba2023towards] found that rising income inequality reduces ecological footprint but exerts a significant drag on economic growth – with critical implications for poverty alleviation in a region with many of the world’s lowest-income countries. In a study of Western African states from 1984 to 2016, Langnel _et al._ [-@langnel2021income] similarly found that income inequality often lowers ecological footprint. The "marginal propensity to emit" hypothesis also found empirical support in Ali [-@ali2023income], which analyzed found a tradeoff relationship between the Gini for income and carbon emissions in 42 middle-income countries between 1990 and 2016. In high-income economies, Hailemariam _et al._ [-@hailemariam2020carbon]’s study of OECD countries found that increasing income inequality (also measured by Gini) was associated with decreasing carbon emissions.

However, there has also been strong empirical evidence for the opposite effect, with lower inequality corresponding to lower environmental impact. In a major early study, Holland _et al._ [-@holland2009cross] found that high income inequality (measured by Gini) was a significant factor predicting anthropogenic biodiversity loss. In a study of 135 countries over the period 1981-2012, Malerba [-@malerba2020trade] found that reducing income inequality consistently reduced the tradeoff between rising prosperity and environmental impact as measured by the "carbon intensity of poverty reduction." In a study of five middle-income ASEAN countries, Masud _et al._ [-@masud2020does] found that from 1985-2015, a tight relationship between income inequality and environmental degradation, and with a bidirectional causality for the bottom 40% of the income distribution. Overall, they concluded that "policymakers should reduce income inequality to strengthen environmental sustainability". Focusing on the relationships between carbon emissions, economic growth, and income inequality before and after the 2008 "Great Recession" across 70 countries, a "robust finding" by Agusdinata _et al._ [-@agusdinata2021economic] was that countries on trajectories of declining emissions were also on trajectories of declining inequality.

The case for the inequality-environment relationship being segmented by development level (i.e., average national income) has also found evidence in analyses of global datasets. Ehigiamusoe _et al._ [-@ehigiamusoe2022nexus] found that while higher income inequality lowers carbon emissions and ecological footprints in high-income countries, it worsens them in middle-income states and the effect of poverty reduction on environmental outcomes seems to dominate in lower-income countries. Similarly, Grunewald [-@Grunewald2017] found that income inequality and carbon emissions were negatively correlated in low- and middle-income countries but positively correlated in high-income countries.

Our findings tend to support the conclusion that lowering economic inequality and lowering environmental impact are synergistic. In particular, in our three-dimensional analyses of country trajectories within the trilemma space, no country started or ended in a region with (relatively) high inequality and high ecological footprint (the upper right quadrant of the cube). And based on their trajectories, all countries first decreased in income inequality before increasing in ecological footprint, regardless of prosperity level or magnitude of displacement from their starting point. This suggests that countries still have developmental "room" to decrease income inequality without increasing environmental impact (Figure 1G; Figure 2C).

Nonetheless, we recognize that the divergent trajectories (including both direction and magnitude of displacement) of countries and pronounced clustering, as well as the large variations in indicator performance in the static analyses, highlight the importance of structural international differences. This applies not only to geographic factors (e.g., population density, natural resource endowments, and climatic conditions determine per-capita biocapacity) but also to past development trajectories and the position of different countries in the structure of the world economy. Figge _et al._ [-@figge2017effects] found that even after controlling for prosperity (average income), globalization significantly increased countries’ ecological footprints globally. And based on analysis of per-capita ecological footprint and Gini for income, Teixidó-Figueras and Duro [-@teixido2015building] concluded that the primary approach to reduce overall global environmental impact is to reduce the large disparities between countries’ per-capita footprints through redistributive policies across countries, prioritizing international equity over economic growth per se. Therefore, structural reforms to international networks of trade and finance – which has come to be increasingly advocated by scholars, activists, policymakers in recent years – could be crucial to overcoming the sustainability trilemma.

Social-ecological feedbacks, which often reinforce the effects of geographic factors and development path dependencies, can also help explain the appearance of "regimes" such as the inequality trap found in the trilemma space. Weinzettel _et al._ [-@Weinzett:2013] found that countries with high per-capita biocapacity (i.e., large natural resource endowments relative to population level) tended to spare more land for nature, forming a virtuous cycle of improving environmental conditions. Similarly, population, environmental carrying capacity, and pollution have been found to have mutually-reinforcing feedbacks across countries [@LaTorr:2019]. This helps to explain the relatively good performance of countries such as Canada, Sweden, and Australia in our analyses. These types of social-ecological feedbacks can create and maintain "red loop" or "green loop" dynamics, entrapping some countries in long-term conditions of high poverty, inequality, and environmental degradation while placing others on a more sustainable trajectory [@Cumming:2014; @Cumming:2018].

#Limitations and Conclusions

Although this study identified new international patterns and trends in the relationships between prosperity, equality, and environmental impact, it still faced some limitations. First, the offshoring of environmental degradation through global supply chains is not fully captured in standard ecological footprint measurements [@VandenBergh:2012]. A consumption-based approach to measuring environmental impacts – as opposed to the more mainstream production-based approach – may help internalize at least some of these externalities (e.g., many resource-intensive goods produced in Asia are consumed in the US and Europe, but the consequent increases in ecological footprint are attributed to the former). However, such an approach is currently constrained by a lack of necessary data, with existing studies tending to focus more narrowly on subsets of commodities and on bilateral or regional trade. If these outsourced impacts were incorporated into the appropriate indicators, our findings would likely show that affluent countries have performed much more poorly with respect to the sustainability trilemma. Addressing this issue has important implications for the international distribution of responsibilities with respect to global commons challenges such as mitigating climate change, biodiversity loss, and pandemic risks - all of which disproportionately affect lower-income countries [@Fankhauser:2014; @Mendelsohn:2006; @IPCC:2021]. It could also lend support for the much-needed reforms to international structures of trade and finance discussed earlier.

We also recognize that there are other valid ways to conceptualize sustainable development. For instance, although they did not account for distributional issues, studies on the relationship between ecological limits, human population, and the economy have a long pedigree [@Ehrlich:1971; @Meadows:1972]. More recently, a popular framework put forward by Raworth [-@RaworthKate:2017] postulated 11 "social thresholds" and 7 planetary boundaries to delimit sustainability. This framework and similar approaches imply equal weight among chosen thresholds, which may mix subjective indicators (e.g., life satisfaction and democratic quality) with objective measures of wellbeing (e.g., nutrition and life expectancy). Such classification schemes are informative but may also be overdetermined: it is conceivable that a different and equally justifiable list of thresholds could be proposed. In addition, there have been calls for analyses to better integrate social and ecological goals to account for the foundational role played by nature [@reyers2020global]. While acknowledging that our "classic" triple bottom line framework is imperfect, we believe it nonetheless captures long-standing and important dimensions of sustainability and how countries have progressed or regressed within the developmental space formed by them.

Whether the synergies and tradeoffs between prosperity, inequality, and the environmental impact identified in this study hold across scales (e.g., at subnational scales or in other analytic units such as cities or corporations), also remains an open question. Furthermore, the feedback mechanisms for interactions among the three dimensions at different scales are also likely to be complex [@Weinzett:2013; @Cumming:2018; @Hamann:2018; @LaTorr:2019]. If there is a unifying concern among countries and across scales, it may be the foundational role of nature, which provides the basic life-support systems for human well-being and the inescapable context for collective development. As our findings strongly suggest, a myopic focus on ever-growing prosperity as the solution to inequality and degradation is misplaced. Indeed, highly prosperous countries do not necessarily have lower levels of inequality, but almost always have higher environmental impacts. One undeniable trend identified in our analysis is that a singular focus on economic growth can undermine progress towards social equality and, most especially, environmental integrity [@arrow2004we; @fanning2022social].

# References {.unnumbered}

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

## List of countries

**Dynamic analysis (N = 59):**

```{r results="asis", echo=FALSE, message=FALSE, warning=FALSE}

load("data/cleaned_trilemma-20230614.RData")
load("data/trilemma_static.Rda")
countries <- readr::read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()

df_dat <- df_dat |> 
    rename(EF = EFconsPerCap, GNI = gni2, Gini = gini)

c <- df_dat |> 
    left_join(countries, by = c("iso3" = "country_code")) |> 
    pull(short_name) |> unique()

cat(c, sep = ", ")
```

**Static analysis (N = 140):**

```{r results="asis", echo=FALSE, message=FALSE, warning=FALSE}
c <- df_all2 |> 
    left_join(countries) |> 
    pull(short_name) |> unique()

cat(c, sep = ", ")
```

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{prosperity_missing_values}
\caption{\textbf{Data quality for prosperity} Different prosperity metrics are compared with respect to their coverage in space (countries) and time (A). Gross national income (GNI) is the metric with higher completion rate (B), with an arbitrary threshold of 30\% missing values, the analysis could include over 100 coutries (C).}
\label{fig:SM_gni_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{inq_missing_values}
\caption{\textbf{Data quality for inequality} Mean Gini income is the most complete statistic of inequality with respect to their coverage in space (countries) and time (A). Alternative metrics such as the Palma index overcome some of the limitations of Gini but have disproportionally more missing values (B). For Gini, the proportion of countries with missing values is above 50\% for any given year. An arbitrary treshold of 30\% threshold for includsion will leave countries like South Africa (country with highers inequality in the world) out of the analysis (D).}
\label{fig:SM_inq_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{ecofp_missing_values}
\caption{\textbf{Data quality for ecological footprint} Coverage of the ecological footprint in space (countries) and time (A) shows that data is relatively complete after mid 1990s (B) An arbitrary threshold of 10 missing values would leave some of the small islands states out of the analysis (C).}
\label{fig:SM_ecofp_missing}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{FigS4}
\caption{\textbf{Trajectories examples} As shown by the experiences of Ukraine and China from 1995 to 2017, historical and structural factors influence the trajectories of countries within the "trilemma space" formed by the economic, social, and environmental dimensions of sustainability: (A) the global financial crisis; (B) the beginning of the Russo-Ukrainian conflict; (C) the accession of China into the World Trade Organization; (D) the beginning of the large-scale anti-corruption crackdown in China.}
\label{fig:S4}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 3.5in]{sm_trajectories}
\caption{\textbf{Trajectories per continent} 59 countries trajectories coloured by continent. Some countries are labelled but not all of them to avoid over plotting.}
\label{fig:S5}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 3.5in]{first_diff}
\caption{\textbf{First difference trajectories} 59 countries trajectories as first difference per variable.}
\label{fig:S6}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{sm_kuznets}
\caption{\textbf{Kuznets hypotheses and our data}  The top panel shows the Kuznets curve (KC) hypothesis (A), the environmental (EKC) counterpart (B), and the logical consequence of a linear relationship if both KC and EKC are true (C). The central panel (D-F) shows a non-linear loess fit on our trajectories data (59 countries, 22 year period), and the bottom panel (G-I) the same fit for the static analysis (140 countries). E,F,H,I deviate completely from the expected hypotheses. D and G show a mild curvature (the coefficient is significant but is 10 orders of magnitude smaller than the linear counterpart, table \ref{tab:1}), and the regions of the curve one can trust less are the extremes that make up the curvature due to the expanding errors. In fact, the far left side of panel D are the first three years of the time series for Spain, Portugal and Italy. The data does not conform to the KC or EKC hypotheses.}
\label{fig:EKC}
\end{figure*}
```

```{r echo=FALSE, results="asis", message=FALSE, warning=FALSE}
#linear regression:
fit <- glm(EF  ~ GNI * Gini + I(GNI^2) + I(Gini^2), 
           family = "gaussian", 
           data = df_dat)

stargazer::stargazer(
    fit, header = FALSE,
    title = "Regression of the trilemma space",
    label = "tab:1")
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 3.5in]{sm_cube}
\caption{\textbf{3D cube of the trilemma space} The surface from regression in Table \ref{tab:1} is shown in gray, trajectory points are show in the three-dimensional space with colors proportional to the environmental footprint axes. The best fit differs from quadratic forms on the KC and EKC axes.}
\label{fig:sm_cube}
\end{figure*}
```

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 4in, height = 4in]{correlogram}
\caption{\textbf{Correlogram} Correlations between the dimensions of the trilemma space}
\label{fig:cors}
\end{figure*}
```